Implementation and comparison of register allocation to translate \lstinline!LLVM--! to x86

### 2023-12-28 article.tex
Before send

Building a graph is relatively straight forward, with the input for every function declared parsed into a tuple type \mintinline{ocaml}{cfg}  of the form:
\begin{center}
\begin{minted}{ocaml}
type cfg = (lbl option * block) * (lbl * block) list
\end{minted}
\end{center}
which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second.

@@

Each function defined in an LLVM program is constructed with the help of

@@

The intermediate representation emitted by the frontend of a compiler serves as a stepping stone independent of the target architecture. The \lstinline!LLVM! infrastructure is the industry standard in terms of bridging this gap and was consequently the library used to translate the semantically annotated abstract syntax tree to executable machine code in the 2022 compilers course. As this project is an expansion on this, it follows naturally to build on this.

@@

Because of this, these instructions are referred to as \textit{terminators}, as their purpose is to disrupt what had otherwise been a linear flow from the beginning of this continus sequence of instructions, henceforth referred to as a \textit{basic block}.

@@

@@

@@

%%% NOTE: Don't include subsections of tthe benchmarks section
\addtocontents{toc}{\protect\setcounter{tocdepth}{1}}

\section{LLVM-- instruction set}
\label{appendix:llvmis}
\begin{enumerate}
    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!,\\\lstinline!srem!, \lstinline!shl!, \lstinline!sub! and \lstinline!xor!
    \item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
    \item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
    \item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint!, \lstinline!trunc!, and \lstinline!zext!
    \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
    \item a  nullary block terminator \lstinline!unreachable!
\end{enumerate}
\end{document}

>>>>

Building a graph is relatively straight forward, with the input for every function declared parsed into a tuple type \mintinline{ocaml}{cfg}  of the form:
\begin{center}
\begin{minted}{ocaml}
type cfg = (lbl option * block) * (lbl * block) list
\end{minted}
\end{center}
which consists of an optionally named entry block in the first part and a list of trailing blocks that are always named in the second. The reason for the first block to be optionally named is the fact that \textit{some} LLVM programs transfer control back to the point of entry, whereas every subsequent block needs to be named because branching can only target it by its name.

@@

%Each function defined in an LLVM program is constructed with the help of

@@

%The intermediate representation emitted by the frontend of a compiler serves as a stepping stone independent of the target architecture. The \lstinline!LLVM! infrastructure is the industry standard in terms of bridging this gap and was consequently the library used to translate the semantically annotated abstract syntax tree to executable machine code in the 2022 compilers course. As this project is an expansion on this, it follows naturally to build on this.

@@

%Because of this, these instructions are referred to as \textit{terminators}, as their purpose is to disrupt what had otherwise been a linear flow from the beginning of this continus sequence of instructions, henceforth referred to as a \textit{basic block}.

\label{appendix:llvmis}
\begin{enumerate}
    \item binary operations \lstinline!add!, \lstinline!and!, \lstinline!ashr!, \lstinline!lshr!, \lstinline!mul!, \lstinline!or!, \lstinline!sdiv!,\\\lstinline!srem!, \lstinline!shl!, \lstinline!sub! and \lstinline!xor!
    \item integer comparison \lstinline!icmp! (conditions being \lstinline!eq!, \lstinline!ne!, \lstinline!sge!, \lstinline!sgt!, \lstinline!sle! and \lstinline!slt!)
    \item memory/address operations \lstinline!alloca!, \lstinline!gep!, \lstinline!load! and \lstinline!store!
    \item  \lstinline!mov! operations \lstinline!gep!, \lstinline!phi!, \lstinline!ptrtoint!, \lstinline!trunc!, and \lstinline!zext!
    \item control flow operations \lstinline!br!, \lstinline!call! and \lstinline!ret!
    \item a  nullary block terminator \lstinline!unreachable!
\end{enumerate}

@@

\subsubsection{Implementing \mintinline{llvm}{call} operation}

The \mintinline{llvm}{call} operation is by far the most complicated one. Mostly because of calling conventions,
The \mintinline{gas}{%rax} register is zeroed before

@@

\subsubsection{\lstinline!benches/fannkuch-redux.ll!}
4 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 172157& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04312x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02806x& \cellcolor[rgb]{0.751953125,0.9990234375,0.75} 1.00936x& \cellcolor[rgb]{0.767578125,0.9990234375,0.75} 1.07378x& \cellcolor[rgb]{0.806640625,0.9990234375,0.75} 1.23096x& \cellcolor[rgb]{0.7998046875,0.9990234375,0.75} 1.20263x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24619x\\
5 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 277849& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.71875x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.74478x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.72175x& \cellcolor[rgb]{0.75,0.9990234375,0.75} 0.74462x& \cellcolor[rgb]{0.7978515625,0.9990234375,0.75} 1.19407x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18339x& \cellcolor[rgb]{0.8017578125,0.9990234375,0.75} 1.20806x\\
6 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 307924& \cellcolor[rgb]{0.8125,0.9990234375,0.75} 1.25112x& \cellcolor[rgb]{0.82421875,0.9990234375,0.75} 1.30051x& \cellcolor[rgb]{0.810546875,0.9990234375,0.75} 1.24357x& \cellcolor[rgb]{0.830078125,0.9990234375,0.75} 1.32219x& \cellcolor[rgb]{0.9990234375,0.9423828125,0.75} 3.80349x& \cellcolor[rgb]{0.9990234375,0.9404296875,0.75} 3.86493x& \cellcolor[rgb]{0.9990234375,0.94140625,0.75} 3.84259x\\
7 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1312806& \cellcolor[rgb]{0.87890625,0.9990234375,0.75} 1.52113x& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62393x& \cellcolor[rgb]{0.8779296875,0.9990234375,0.75} 1.51710x& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62624x& \cellcolor[rgb]{0.9990234375,0.86328125,0.75} 6.35235x& \cellcolor[rgb]{0.9990234375,0.8515625,0.75} 6.70982x& \cellcolor[rgb]{0.9990234375,0.8642578125,0.75} 6.31152x\\
8 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 11168259& \cellcolor[rgb]{0.900390625,0.9990234375,0.75} 1.60627x& \cellcolor[rgb]{0.9189453125,0.9990234375,0.75} 1.67856x& \cellcolor[rgb]{0.947265625,0.9990234375,0.75} 1.79351x& \cellcolor[rgb]{0.9140625,0.9990234375,0.75} 1.66070x& \cellcolor[rgb]{0.9990234375,0.8349609375,0.75} 7.25991x& \cellcolor[rgb]{0.9990234375,0.833984375,0.75} 7.29830x& \cellcolor[rgb]{0.9990234375,0.83984375,0.75} 7.09878x\\
9 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 115754944& \cellcolor[rgb]{0.900390625,0.9990234375,0.75} 1.60476x& \cellcolor[rgb]{0.9296875,0.9990234375,0.75} 1.72352x& \cellcolor[rgb]{0.9052734375,0.9990234375,0.75} 1.62536x& \cellcolor[rgb]{0.9208984375,0.9990234375,0.75} 1.68769x& \cellcolor[rgb]{0.9990234375,0.828125,0.75} 7.46779x& \cellcolor[rgb]{0.9990234375,0.8173828125,0.75} 7.81576x& \cellcolor[rgb]{0.9990234375,0.83203125,0.75} 7.36417x\\
10 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1352747472& \cellcolor[rgb]{0.8857421875,0.9990234375,0.75} 1.54700x& \cellcolor[rgb]{0.935546875,0.9990234375,0.75} 1.74875x& \cellcolor[rgb]{0.89453125,0.9990234375,0.75} 1.58381x& \cellcolor[rgb]{0.9169921875,0.9990234375,0.75} 1.67440x& \cellcolor[rgb]{0.9990234375,0.818359375,0.75} 7.77269x& \cellcolor[rgb]{0.9990234375,0.8193359375,0.75} 7.74449x& \cellcolor[rgb]{0.9990234375,0.8173828125,0.75} 7.82170x\\
11 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 16942071879& \cellcolor[rgb]{0.88671875,0.9990234375,0.75} 1.55031x& \cellcolor[rgb]{0.9267578125,0.9990234375,0.75} 1.71033x& \cellcolor[rgb]{0.8876953125,0.9990234375,0.75} 1.55567x& \cellcolor[rgb]{0.9228515625,0.9990234375,0.75} 1.69492x& \cellcolor[rgb]{0.9990234375,0.814453125,0.75} 7.92913x& \cellcolor[rgb]{0.9990234375,0.8017578125,0.75} 8.32639x& \cellcolor[rgb]{0.9990234375,0.8212890625,0.75} 7.70198x\\
12 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 235117543385& \cellcolor[rgb]{0.8974609375,0.9990234375,0.75} 1.59287x& \cellcolor[rgb]{0.923828125,0.9990234375,0.75} 1.70029x& \cellcolor[rgb]{0.8798828125,0.9990234375,0.75} 1.52228x& \cellcolor[rgb]{0.927734375,0.9990234375,0.75} 1.71456x& \cellcolor[rgb]{0.9990234375,0.8095703125,0.75} 8.05498x& \cellcolor[rgb]{0.9990234375,0.8095703125,0.75} 8.08426x& \cellcolor[rgb]{0.9990234375,0.806640625,0.75} 8.17452x\\
\caption{\label{tab:fannkuch}Benchmark of \texttt{benches/sieven.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}

@@

%%% NOTE: Don't include subsections of tthe benchmarks section
%\addtocontents{toc}{\protect\setcounter{tocdepth}{1}}

\end{document}
